Stabilization of carbon in mineral soils from mangroves of the Sin\'{u} river delta, Colombia

### 2017-12-12 Manuscript/carbonStabilizationMangroves.tex
Fixed issue with LaTeX preamble
Minor changes based on pdf provided by reviewer 2

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended, referee]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{here}
\usepackage{upgreek}
\usepackage{wasysym}
\usepackage{natbib}
\usepackage{lineno}
\usepackage{url}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
\journalname{Wetlands Ecology and Management}
%
\begin{document}

@@

\section{Introduction}
\label{intro}
Although mangroves are ecosystems with some of the largest levels of carbon storage on earth \citep{Donato2011, Alongi2012}, there are  large variations on the amount of C stored in various systems, particularly for soils. For instance, mangrove ecosystems dominated by \emph{Rhizophora} spp. in Peninsular Malaysia store between 479 to 2205 Mg C ha$^{-1}$ in the belowground and soil pool \citep{Alongi2012}, and similar levels of variability have been observed at other sites \citep{Jardine2014}. It is unclear however, what are the main determinants of observed differences in soil C storage across diverse mangrove systems. It is possible that differences in soil carbon storage are due to differences in the level of productivity of different mangrove systems, or due to other external sources such as sediment transport.

@@

\section{Materials and methods} \label{sec:1}
\subsection{Study site}
The study site is located on the northwestern Caribbean coast in Colombia (9$^{\circ}$23'N 75$^{\circ}$52'W), which is part of the southern extreme of Morrosquillo gulf and it is locally known as Cispat\'a bay. The coastal zone is characterized by the Sin\'{u} river delta and a complex estuarine lagoon system. Extensive wetlands and mangroves dominate this area. The Sin\'{u} river has its origin in the northern part of the western Andean mountains, and between 1938 and 1945 changed its course creating a new delta \citep{suarez2004}. The current estuarine mangrove system was established in the previous river delta. 
The area includes mainly 3 mangrove species. While fringe forests are dominated by the species \textit{Rhizophora mangle}, basin forests are dominated by \textit{Avicennia germinans}. \textit{Laguncularia racemosa} occurs in both forest types.

@@

\subsection{Laboratory analyses}
After collection, samples were oven-dried at 70$^{\circ}$C for 5 days. Samples for bulk density measurements were oven dried at 105$^{\circ}$C. We calculated soil bulk density as dry mass divided by fresh volume, which matches the used sampling rings \mbox{(V = 98.52 cm$^3$)}. Each sample was ground for 3 min at a frequency of 25 Hz using a Retsch MM 400 ball mill.

@@

There were no differences between basin and fringe mangroves in terms of their composition of the mineral fraction (Table \ref{tab:2}). The proportion of mineral soil is bigger in basin than in fringe mangroves. Fringe mangrove soils show in contrast a higher proportion of halite than basin soils.

@@

The oxalate extraction dissolved much of the iron and aluminum from the amorphous materials, whereas the dithionite extraction dissolved only little of the crystalline iron oxides as well as little of the amorphous materials \mbox{(Fig. \ref{fig:6})}. Oxalate- and dithionite extracted Al showed significant differences among fringe and mangrove forest soils ($p$-value $< 0.001$, $F = 16.35$, 39 d.f.). Similarly, oxalate and dithionite extracted Fe showed significant differences between both soils ($p-$value $< 0.001$, $F = 33.45$, 39 d.f.).
Basin soil samples had two times higher concentrations of Al (Al$_d$+Al$_o$) than fringe soils, and five times higher Fe values. While riverine samples had similar Al and Fe concentrations than those of basin mangroves in all 4 extraction patterns, metal contents of sand samples were constantly low. Fe$_o$ contents differed extremely for basin soils, especially for the first 2 depth intervals. Both basin and fringe soils showed an increase of metal oxides within the first depth intervals and in turn a decrease within the deeper layers.

@@

Higher TOC values in basin mangrove soils reflect higher rates of organic matter accumulation. According to  \citet{Bolivar2015}, the percentage of clay is similar between both mangrove types. However, the silt fraction dominates in all soil profiles in basin mangroves, while sand dominates in fringe mangrove soils. It has been well established that soil particles with greater surface area, as typical of finer textures like those found in basin mangroves, deteriorate drainage conditions which in turn increase retention of organic matter \citep{prasad2008}. Because TOC contents are linked to \%OC, we found also an increase of TOC in depth levels of 80-100 cm (basin) and 60-80 cm (fringe).

>>>>

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
%\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
%gsave
%newpath
%  20 20 moveto
%  20 220 lineto
%  220 220 lineto
%  220 20 lineto
%closepath
%2 setlinewidth
%gsave
%  .4 setgray fill
%grestore
%stroke
%grestore
%\end{filecontents*}
%%
%\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended, referee]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{here}
\usepackage{upgreek}
\usepackage{wasysym}
\usepackage{natbib}
\usepackage{lineno}
\usepackage{url}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
\journalname{Wetlands Ecology and Management}
%
\begin{document}

@@

\section{Introduction}
\label{intro}
Although mangroves are ecosystems with some of the largest levels of carbon storage on earth \citep{Donato2011, Alongi2012}, there are  large variations on the amount of C stored in various systems, particularly soils. For instance, mangrove ecosystems dominated by \emph{Rhizophora} spp. in Peninsular Malaysia store between 479 to 2205 Mg C ha$^{-1}$ in the belowground and soil pool \citep{Alongi2012}, and similar levels of variability have been observed at other sites \citep{Jardine2014}. It is unclear however, what are the main determinants of observed differences in soil C storage across diverse mangrove systems. It is possible that differences in soil carbon storage are due to differences in the level of productivity of different mangrove systems, or due to other external sources such as sediment transport.

@@

\section{Materials and methods} \label{sec:1}
\subsection{Study site}
The study site is located on the northwestern Caribbean coast in Colombia (9$^{\circ}$23'N 75$^{\circ}$52'W), which is part of the southern extreme of Morrosquillo gulf and it is locally known as Cispat\'a bay. The coastal zone is characterized by the Sin\'{u} river delta and a complex estuarine lagoon system that covers approximately 5,098 ha. Extensive wetlands and mangroves dominate this area. The Sin\'{u} river has its origin in the northern part of the western Andean mountains, and between 1938 and 1945 changed its course creating a new delta \citep{suarez2004}. The current estuarine mangrove system was established in the previous river delta. 
Fringe forests are dominated by  \textit{Rhizophora mangle}, basin forests are dominated by \textit{Avicennia germinans}. \textit{Laguncularia racemosa} occurs in both forest types.

@@

\subsection{Laboratory analyses}
After collection, samples were oven-dried at 70$^{\circ}$C for 5 days. Samples for bulk density measurements were oven dried at 105$^{\circ}$C. We calculated soil bulk density as dry mass divided by fresh volume \mbox{(V = 98.52 cm$^3$)}. Each sample was ground for 3 min at a frequency of 25 Hz using a Retsch MM 400 ball mill.

@@

There were no differences between basin and fringe mangroves in terms of their composition of the mineral fraction (Table \ref{tab:2}). The proportion of mineral soil is larger in basin than in fringe mangroves. Fringe mangrove soils show in contrast a higher proportion of halite than basin soils.

@@

The oxalate extraction dissolved much of the poorly crystalline Fe and Al oxides from the amorphous materials, whereas the dithionite extraction dissolved the crystalline Fe oxides as well as the amorphous materials \mbox{(Fig. \ref{fig:6})}. Oxalate- and dithionite extracted Al showed significant differences among fringe and mangrove forest soils ($p$-value $< 0.001$, $F = 16.35$, 39 d.f.). Similarly, oxalate and dithionite extracted Fe showed significant differences between both soils ($p-$value $< 0.001$, $F = 33.45$, 39 d.f.).
Basin soil samples had two times higher concentrations of Al (Al$_d$+Al$_o$) than fringe soils, and five times higher Fe values. While riverine samples had similar Al and Fe concentrations than those of basin mangroves in all 4 extraction patterns, metal contents of sand samples were constantly low. Fe$_o$ contents differed extremely for basin soils, especially for the first 2 depth intervals. Both basin and fringe soils showed an increase of metal oxides within the first depth intervals and in turn a decrease within the deeper layers.

@@

Higher TOC values in basin mangrove soils reflect higher rates of organic matter accumulation. According to  \citet{Bolivar2015}, the percentage of clay is similar between both mangrove types. However, the silt fraction dominates in all soil profiles in basin mangroves, while sand dominates in fringe mangrove soils. It has been well established that soil particles with greater surface area, as typical of finer textures like those found in basin mangroves, decrease drainage and decomposition of organic matter \citep{prasad2008}. Because TOC contents are linked to \%OC, we found also an increase of TOC in depth levels of 80-100 cm (basin) and 60-80 cm (fringe).
